---
title: "The Meyers Effect"
subtitle: "Scraping and Tidying"
format: html
editor: visual
---

## Loading Libraries\`

```{r}
# Load packages needed for scraping, tidying, and iteration
library(tidyverse)
library(stringr)
library(httr2)
library(httr)
library(purrr)
library(rvest)
library(readr)
library(polite)
library(jsonlite)
library(dplyr)
```

## Check Scraping Permissions

Before scraping any site, we verify that robots.txt allows access.

```{r}
# Confirm scraping is permitted on our sources
robotstxt::paths_allowed("https://www.imdb.com")
robotstxt::paths_allowed("https://en.wikipedia.org")

# Create a data directory to cache raw scraped files
dir.create("data", showWarnings = FALSE)
```

## Define Film Lookup Tables

We define two lookup tibbles — one for Meyers' full filmography (as writer, director, and/or producer), and one for our comparison rom-coms (chosen based on famous movies that compitted with Meyers' back then). These act as our "key" tables.

```{r}
# imdb_id: the tt-code from each film's IMDb URL
# bom_slug: the path segment used in Box Office Mojo URLs
# rt_slug:  the path segment used in Rotten Tomatoes URLs
meyers_films <- tibble(
  title        = c(
    "Private Benjamin",
    "Irreconcilable Differences",
    "Baby Boom",
    "Father of the Bride",
    "I Love Trouble",
    "Father of the Bride Part II",
    "The Parent Trap",
    "What Women Want",
    "Something's Gotta Give",
    "The Holiday",
    "It's Complicated",
    "The Intern"
  ),
  year         = c(1980, 1984, 1987, 1991, 1994, 1995, 1998, 2000, 2003, 2006, 2009, 2015),
  meyers_role  = c(
    "Writer, Producer",
    "Writer, Producer",
    "Writer, Producer",
    "Writer, Producer",
    "Writer, Producer",
    "Writer, Director, Producer",
    "Writer, Director, Producer",
    "Director, Producer",
    "Writer, Director, Producer",
    "Writer, Director, Producer",
    "Writer, Director, Producer",
    "Writer, Director, Producer"
  ),
  imdb_id      = c(
    "tt0080373", "tt0087564", "tt0092699", "tt0101862",
    "tt0110278", "tt0114692", "tt0120783", "tt0207201",
    "tt0337741", "tt0457939", "tt1178663", "tt2234222"
  ),
  bom_slug     = c(
    "private-benjamin", "irreconcilable-differences", "baby-boom",
    "father-of-the-bride", "i-love-trouble", "father-of-the-bride-part-ii",
    "the-parent-trap", "what-women-want", "somethings-gotta-give",
    "the-holiday", "its-complicated", "the-intern"
  ),
  rt_slug      = c(
    "private_benjamin", "irreconcilable_differences", "baby_boom",
    "father_of_the_bride", "i_love_trouble", "father_of_the_bride_part_ii",
    "the_parent_trap", "what_women_want", "somethings_gotta_give",
    "the_holiday", "its_complicated", "the_intern"
  ),
  group        = "Meyers"
)

#Comparison rom-coms lookup 
comparison_films <- tibble(
  title        = c(
    "Notting Hill",
    "Bridget Jones's Diary",
    "My Big Fat Greek Wedding",
    "How to Lose a Guy in 10 Days",
    "13 Going on 30",
    "Hitch",
    "The Devil Wears Prada",
    "Knocked Up",
    "Mamma Mia!",
    "The Proposal",
    "Crazy, Stupid, Love",
    "About Time",
    "The Other Woman",
    "Trainwreck"
  ),
  year         = c(1999, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2011, 2013, 2014, 2015),
  meyers_role  = NA_character_,
  imdb_id      = c(
    "tt0125439", "tt0243155", "tt0259446", "tt0343660",
    "tt0337563", "tt0386588", "tt0458352", "tt0478311",
    "tt0382932", "tt1041829", "tt1570728", "tt2219695",
    "tt2203939", "tt3153910"
  ),
  bom_slug     = c(
    "notting-hill", "bridget-joness-diary", "my-big-fat-greek-wedding",
    "how-to-lose-a-guy-in-10-days", "13-going-on-30", "hitch",
    "the-devil-wears-prada", "knocked-up", "mamma-mia",
    "the-proposal", "crazy-stupid-love", "about-time",
    "the-other-woman", "trainwreck"
  ),
  rt_slug      = c(
    "notting_hill", "bridget_joness_diary", "my_big_fat_greek_wedding",
    "how_to_lose_a_guy_in_10_days", "13_going_on_30", "hitch",
    "the_devil_wears_prada", "knocked_up", "mamma_mia",
    "the_proposal", "crazy_stupid_love", "about_time",
    "the_other_woman", "trainwreck"
  ),
  group        = "Comparison"
)

# Combine into one master lookup 
all_films <- bind_rows(meyers_films, comparison_films)

all_films
```

## Scrape IMDb Metadata

```{r}
omdb_key <- "d766c35e"

fetch_omdb <- function(imdb_id, title) {
  resp <- request("https://www.omdbapi.com/") |>
    req_url_query(i = imdb_id, apikey = omdb_key) |>
    req_perform() |>
    resp_body_json()

  if (resp$Response == "False") return(NULL)

  tibble(
    imdb_id      = imdb_id,
    title        = title,
    imdb_rating  = parse_number(resp$imdbRating),
    vote_count   = parse_number(resp$imdbVotes),
    runtime_mins = parse_number(resp$Runtime),
    mpaa_rating  = resp$Rated,
    genres       = resp$Genre
  )
}

imdb_data <- map2(all_films$imdb_id, all_films$title, fetch_omdb) |>
  bind_rows()

```

## Scrape Financial Data from Wikipedia

Since Box Office Mojo and Rotten Tomatoes both render their content via JavaScript, which means `rvest` receives a skeleton page with no actual data in it, we use Wikipedia's film infoboxes that are plain static HTML and contain the same financial figures drawn from the same primary sources, and works better with `rvest`.

Wikipedia infoboxes are two-column HTML tables where each row is a `<tr>` containing a `<th>` (label) and a `<td>` (value).

```{r}
# Walks each <tr> in the infobox, checks if its <th> matches our pattern,
# and returns the paired <td> text. This is alignment-safe regardless of
# how many image/caption rows appear in the infobox.
get_infobox_value <- function(infobox, pattern) {
  rows <- infobox |> html_elements("tr")
  for (row in rows) {
    label <- row |> html_element("th") |> html_text2()
    if (!is.na(label) && str_detect(label, regex(pattern, ignore_case = TRUE))) {
      value <- row |> html_element("td") |> html_text2()
      return(value)
    }
  }
  return(NA_character_)
}

# scrape one Wikipedia film infobox
scrape_wiki_financials <- function(wiki_slug, title) {
  url     <- str_c("https://en.wikipedia.org/wiki/", wiki_slug)
  session <- bow(url, user_agent = "Educational scraping project / DS 334")
  page    <- scrape(session)
  
  # Some articles use .infobox-film, others just .infobox
  infobox <- page |> html_element(".infobox-film, .infobox")
  
  # Guard: if no infobox found at all, return NAs rather than crash map2
  if (is.null(infobox) || inherits(infobox, "xml_missing")) {
    warning(str_c("No infobox found for: ", title, " (slug: ", wiki_slug, ")"))
    return(tibble(
      wiki_slug = wiki_slug, title = title,
      budget_raw = NA_character_, box_office_raw = NA_character_,
      opening_raw = NA_character_, distributor = NA_character_
    ))
  }
  
  tibble(
    wiki_slug      = wiki_slug,
    title          = title,
    budget_raw     = get_infobox_value(infobox, "Budget"),
    box_office_raw = get_infobox_value(infobox, "Box office"),
    opening_raw    = get_infobox_value(infobox, "Opening"),
    distributor    = get_infobox_value(infobox, "Distributed by|Distribution")
  )
}

# Add Wikipedia slugs to the lookup table 
# Slugs must match the exact Wikipedia article title (underscores, disambiguation).
# Verify each slug by visiting en.wikipedia.org/wiki/<slug> in a browser if unsure.
# Note: apostrophes in URLs should be encoded as %27
meyers_films <- meyers_films |>
  mutate(wiki_slug = c(
    "Private_Benjamin",
    "Irreconcilable_Differences_(film)",
    "Baby_Boom_(film)",
    "Father_of_the_Bride_(1991_film)",
    "I_Love_Trouble",
    "Father_of_the_Bride_Part_II",
    "The_Parent_Trap_(1998_film)",
    "What_Women_Want_(film)",
    "Something%27s_Gotta_Give_(film)",
    "The_Holiday",
    "It%27s_Complicated_(film)",
    "The_Intern_(2015_film)"
  ))

comparison_films <- comparison_films |>
  mutate(wiki_slug = c(
    "Notting_Hill_(film)",
    "Bridget_Jones%27s_Diary_(film)",
    "My_Big_Fat_Greek_Wedding",
    "How_to_Lose_a_Guy_in_10_Days_(film)",
    "13_Going_on_30",
    "Hitch_(film)",
    "The_Devil_Wears_Prada_(film)",
    "Knocked_Up",
    "Mamma_Mia!_(film)",
    "The_Proposal_(2009_film)",
    "Crazy,_Stupid,_Love.",        
    "About_Time_(film)",
    "The_Other_Woman_(2014_film)",
    "Trainwreck_(film)"
  ))

# Rebuild master lookup with wiki_slug included
all_films <- bind_rows(meyers_films, comparison_films)

# Iterate over all films 

safe_scrape_wiki <- possibly(scrape_wiki_financials, otherwise = tibble(
  wiki_slug = NA_character_, title = NA_character_,
  budget_raw = NA_character_, box_office_raw = NA_character_,
  opening_raw = NA_character_, distributor = NA_character_
))

wiki_raw  <- map2(all_films$wiki_slug, all_films$title, safe_scrape_wiki)
wiki_data <- bind_rows(wiki_raw)

#  check which films are missing financial data?
wiki_data |> filter(is.na(budget_raw) | is.na(box_office_raw)) |> select(title, wiki_slug)

wiki_data
```

## Tidy and Merge All Sources

We use `left_join` to combine all scraped sources with our master lookup table. Dollar strings are cleaned with `parse_number()` .

```{r}
# clean financial columns from Wikipedia infoboxes 
wiki_clean <- wiki_data |>
  mutate(across(c(budget_raw, box_office_raw, opening_raw),
                ~ str_remove_all(.x, "\\[\\d+\\]"))) |>
  mutate(
    budget_usd     = parse_number(budget_raw),
    box_office_usd = parse_number(box_office_raw),
    opening_usd    = parse_number(opening_raw),
    budget_usd     = if_else(budget_usd < 10000, budget_usd * 1e6, budget_usd),
    box_office_usd = if_else(box_office_usd < 10000, box_office_usd * 1e6, box_office_usd)
  ) |>
  select(wiki_slug, title, budget_usd, box_office_usd, opening_usd, distributor)

#  rename vote_count 
imdb_clean <- imdb_data |>
  rename(vote_count_num = vote_count)

#  join IMDb and Wikipedia data to the master lookup 
films_tbl <- all_films |>
  left_join(imdb_clean, by = c("imdb_id", "title")) |>
  left_join(wiki_clean, by = c("wiki_slug", "title"))

#  derive computed variables 
films_tbl <- films_tbl |>
  mutate(
    profit_margin_pct = ((box_office_usd - budget_usd) / budget_usd) * 100,
    roi_category = case_when(
      profit_margin_pct >= 50 ~ "Hit",
      profit_margin_pct >= 0  ~ "Break-even",
      profit_margin_pct <  0  ~ "Flop",
      TRUE                    ~ NA_character_
    ),
    meyers_directed = str_detect(replace_na(meyers_role, ""), "Director")
  )

films_tbl
```

## Build Supplementary Tibbles

Some variables are not one-to-one with films. We unnested them into separate tidy tibbles for use in specific visualizations.

```{r}
# genre_tbl
genre_tbl <- films_tbl |>
  select(title, year, group, genres) |>
  filter(!is.na(genres)) |>
  mutate(genre = str_split(genres, ",\\s*")) |>
  unnest(genre) |>
  select(-genres)

genre_tbl

# role_tbl
role_tbl <- films_tbl |>
  filter(group == "Meyers") |>
  select(title, year, meyers_role) |>
  mutate(role = str_split(meyers_role, ",\\s*")) |>
  unnest(role) |>
  mutate(role = str_trim(role)) |>
  select(-meyers_role)

role_tbl
```

```{r}

write_csv(films_tbl,  "data/films_tbl.csv")
write_csv(genre_tbl,  "data/genre_tbl.csv")
write_csv(role_tbl,   "data/role_tbl.csv")
write_csv(imdb_clean,  "data/imdb_data.csv")
write_csv(wiki_clean,  "data/wiki_data.csv")

```
