---
title: "The Nancy Meyers effect"
subtitle: "Analysing data from Holywood's famous rom-coms"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    theme:            #We want to get a movie themed aesthetic
      bg: "#F5F0E8"          
      fg: "#1a1a1a"          
      primary: "#1a1a1a"     
      secondary: "#8B0000"   
      base_font: "Libre Baskerville"    
      heading_font: "Playfair Display" 
      code_font: "Courier Prime"        

    vertical_layout: scroll
runtime: shiny
---

```{r setup, include=FALSE}
# Load packages needed for scraping, tidying, and iteration
library(tidyverse)
library(stringr)
library(httr2)
library(httr)
library(purrr)
library(rvest)
library(readr)
library(polite)
library(jsonlite)
library(dplyr)
library(shiny)
library(ggplot2)
library(plotly)
library(DT)
library(flexdashboard)
library(ggrepel)
library(here)
library(base64enc)



films_tbl <- read_csv("data/films_tbl.csv")
genre_tbl <- read_csv("data/genre_tbl.csv")

films_tbl <- films_tbl |>
  mutate(
    profit_margin_pct = ((box_office_usd - budget_usd) / budget_usd) * 100,
    decade = floor(year / 10) * 10
  )
```



About
=====================================

Column
-------------------------------------

### Background

**Who is Nancy Meyers? Why are we so invested in her?**

Nancy Meyers is one of Hollywood's most successful filmmakers, and one of its most underrated. She was born on December 8, 1949, in Philadelphia, Pennsylvania. She started her career in the 1970s writing screenplays, and went on to write, direct, and produce some of the most beloved romantic comedies of the past four decades.

Her films include **The Parent Trap** (1998), **Something's Gotta Give** (2003), **The Holiday** (2006), **It's Complicated** (2009), and **The Intern** (2015). They are funny, warm, and incredibly well-made. They also made a lot of money. Every film she has directed turned a profit record very few directors can claim.


### About this project

We will looks at all 12 films Nancy Meyers was involved in between 1980 and 2015, and compares them against 14 other rom-coms from the same era. We explore four dimensions of her career 

**1. Financial Performance**
Do Meyers films consistently make more money relative to their budget than other rom-coms? We measure this using ROI (worldwide gross divided by production budget), to see whether her commercial success is a pattern or just luck.

**2. Budget vs. Box Office**
How much did each film cost to make, and how much did it earn? We plot production budget against worldwide box office to see which films made the most money relative to what was spent. Films above the diagonal line earned more than they cost, and we want to know if Meyers films consistently land above that line compared to other rom-coms.

**3. Career Evolution**
How did Nancy Meyers grow as a filmmaker since the start of her career? We trace how her budgets, box office returns, and IMDb ratings changed over time, and whether taking on the director's chair made a difference.

**4. Genre Patterns**
What makes a Meyers film a Meyers film? We look at genre tags across all 26 films to see whether her work stands apart from the comparison set and whether those differences might explain why audiences see her so differently.

We chose these four dimensions because together they tell the full story of why she matters, not just as a filmmaker, but as a case study in how Hollywood measures success. Box office numbers tell us what audiences voted for with their wallets. Budgets tell us how much the studios trusted her. Ratings tell us what critics thought versus what real viewers felt. And genre patterns tell us what kind of stories she chose to tell. Taken together, they make the case that Nancy Meyers has been quietly doing something remarkable for 35 years, and that the data backs it up.

**Data sources:**

  [Wikipedia](https://en.wikipedia.org) — financial data
  [IMDb](https://www.imdb.com) — ratings, runtime, genres



ROI by Film
=====================================

Column {.sidebar}
-------------------------------------

### Filter

```{r}
selectInput("decade_filter", "Decade",
            choices = c("All", "1980s", "1990s", "2000s", "2010s"),
            selected = "All")
```


### About this graph

Profit margin is one of the most honest metrics in Hollywood. Unlike raw 
box office numbers, which can be inflated by massive budgets, profit margin 
tells us how efficiently a film converted its investment into returns. A 
blockbuster that cost $300 million and earned $400 million is actually less 
impressive than a $30 million film that earned $150 million. This is exactly 
why comparing Nancy Meyers' films against similarly popular films of her era 
matters. Meyers built her career on mid-budget romantic comedies at a time 
when studios were increasingly betting on expensive franchises and action 
spectacles. If her films consistently outperformed comparable titles in profit 
margin, that tells us something important: her formula, witty dialogue, 
relatable adult characters, aspirational settings, was not just critically 
appreciated but financially shrewd. The comparison group here includes films 
that were commercially successful and culturally visible during the same 
decades, giving us a fair baseline. By filtering across decades, you can also 
trace whether her financial efficiency held steady over time or shifted as 
Hollywood's economic landscape changed around her. The dashed line marks 
break-even , anything to the right was profitable.


Column
-------------------------------------

### Profit Margin by Film

```{r}
renderPlotly({
  p <- films_tbl |>
    filter(!is.na(profit_margin_pct)) |>
    filter(input$decade_filter == "All" |
             decade == as.numeric(substr(input$decade_filter, 1, 4))) |>
    mutate(title = fct_reorder(title, profit_margin_pct)) |>
    ggplot(aes(x = profit_margin_pct, y = title, fill = group,
                   text = paste("Title:", title,
                        "<br>Margin:", round(profit_margin_pct, 1), "%",
                        "<br>Group:", group))) +
    geom_col() +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
    scale_fill_manual(values = c("Meyers" = "#C9A84C", "Comparison" = "#4A2C6E")) +
    labs(title = "Profit Margin by Film",
         x = "Profit Margin %", y = NULL, fill = NULL) +
    theme_minimal()
  
  ggplotly(p, tooltip = "text")
    
})
```

### Caption

*This chart shows the **profit margin** for each film, calculated as
`(box office - budget) / budget × 100`. Purple bars are Nancy Meyers
films; teal bars are comparison films. Films to the right of the
dashed line turned a profit.*



Budget vs Box Office
=====================================

Column {.sidebar}
-------------------------------------

### Filter

```{r}
selectInput("highlight", "Highlight a film",
            choices = c("None", sort(films_tbl$title)))
```

### About this graph

Raw box office numbers only tell half the story. A film that earns $200 
million sounds impressive until you learn it cost $250 million to make. 
This scatter plot puts both numbers on the same canvas, letting us see not 
just how much each film earned, but how that earning relates to what was 
spent. The diagonal dashed line is the break-even threshold, films plotted 
above it turned a profit, films below it lost money. Both axes use a 
logarithmic scale, which prevents a handful of massive blockbusters from 
compressing everything else into an unreadable cluster. This view is 
particularly revealing for Nancy Meyers because her films occupy a specific 
niche: they are not cheap independent productions, but they are also not 
nine-figure franchise tentpoles. They sit in the mid-budget space that 
Hollywood has increasingly abandoned in favor of extremes. Seeing where her 
films land relative to the comparison group, and how consistently they 
appear above the break-even line, helps us understand whether her budgetary 
approach was strategically sound. Use the dropdown to highlight any specific 
film and see exactly where it falls in the landscape of its peers.

Column
-------------------------------------

### Budget vs. Box Office Bubble Chart

```{r}
renderPlotly({
  p2 <- films_tbl |>
    filter(!is.na(budget_usd), !is.na(box_office_usd)) |>
    mutate(label = if_else(title == input$highlight, title, NA_character_)) |>
    ggplot(aes(x = budget_usd, y = box_office_usd, size = vote_count_num, color = group,
                text = paste("Title:", title,
                        "<br>Budget:", scales::dollar(budget_usd, scale = 1e-6, suffix = "M"), 
                        "<br>Group:", group))) +
    geom_abline(linetype = "dashed", color = "gray50") +
    geom_point(alpha = 0.7) +
    geom_text(aes(label = label), hjust = -0.1, size = 4, show.legend = FALSE) +
    scale_x_log10(labels = scales::label_dollar(scale = 1e-6, suffix = "M")) +
    scale_y_log10(labels = scales::label_dollar(scale = 1e-6, suffix = "M")) +
    scale_color_manual(values = c("Meyers" = "#C9A84C", "Comparison" = "#4A2C6E")) +
    labs(title = "Budget vs Box Office",
         x = "Production Budget", y = "Worldwide Box Office",
         color = NULL, size = "Vote count") +
    theme_minimal()
  ggplotly(p2, tooltip = "text")
})
```

### Caption
*Each bubble represents one film. The dashed line marks the
**break-even point** — films above it earned more than they cost.
Bubble size reflects IMDb vote count. Use the dropdown to label
a specific film.*



Career Timeline
=====================================

Column {.sidebar}
-------------------------------------

### Filter

```{r}
radioButtons("timeline_y", "Y Axis",
             choices = c("Budget"      = "budget_usd",
                         "Box Office"  = "box_office_usd",
                         "IMDb Rating" = "imdb_rating"),
             selected = "budget_usd",
             inline = TRUE)
```

### About this graph 

A filmmaker's career is rarely a straight line. Budgets grow and shrink, 
critical reception shifts, and roles evolve. This timeline traces Nancy 
Meyers' trajectory across four decades, letting you toggle between three 
different lenses: production budget, worldwide box office, and IMDb rating. 
Each offers a different perspective on her career arc. Budget tells us how 
much trust studios placed in her at each stage. Box office tells us how 
audiences responded. IMDb rating captures a longer-term cultural verdict, 
reflecting how her films have aged in public memory. The color distinction 
between her directing credits and her writer/producer roles is intentional,
it lets us ask whether she performed differently in different capacities. Did 
her budgets increase when she took the director's chair? Did her box office 
hold up when she stepped back to produce? These are questions that matter 
when evaluating a career that spanned multiple industry eras, from the 
pre-internet Hollywood of the 1980s to the streaming-disrupted landscape of 
the 2010s. The gray connecting line is deliberately neutral, it simply 
guides the eye through chronology, while the colored points carry the 
interpretive weight.

Column
-------------------------------------

### Nancy Meyers Career Timeline

```{r}
renderPlotly({
  y_var <- input$timeline_y

 p3 <- films_tbl |>
    filter(group == "Meyers", !is.na(.data[[y_var]])) |>
    mutate(role_label = if_else(meyers_directed, "Director", "Writer / Producer")) |>
    ggplot(aes(x = year, y = .data[[y_var]], color = role_label,
                text = paste("Title:", title,
                        "<br>Year:", year,  
                        "<br>Role:", role_label))) +
    geom_line(aes(group = 1), color = "gray80") +
    geom_point(size = 4) +
    scale_color_manual(values = c("Director" = "#4A2C6E", "Writer / Producer" = "#B5858A")) +
    scale_y_log10(labels = scales::label_dollar(scale = 1e-6, suffix = "M")) +
    labs(title = "Nancy Meyers Career Timeline",
         x = "Year", y = y_var, color = NULL) +
    theme_minimal()
   ggplotly(p3, tooltip = "text")

})
```

### Caption
*Traces Nancy Meyers' career from the 1980s to 2010s. Points are
colored by her **role** on each film — whether she directed it or
served as writer/producer only.*



Genre Breakdown
=====================================

Column {.sidebar}
-------------------------------------

### Filter

```{r}
checkboxGroupInput("genre_filter", "Genres",
                   choices  = sort(unique(genre_tbl$genre)),
                   selected = sort(unique(genre_tbl$genre)),
                   inline   = TRUE)
```

### About this graph

Genre is the language Hollywood uses to market films and manage audience 
expectations, but it is also a meaningful analytical category. Nancy Meyers 
is most commonly associated with romantic comedy, but her filmography is more 
varied than that label suggests. This chart examines the full genre 
composition of her work and compares it directly to the films in our 
comparison group. The comparison group was selected to reflect films that were 
commercially and culturally prominent during the same period, so differences 
in genre distribution are meaningful rather than incidental. If Meyers 
concentrated heavily in genres that her peers avoided, that tells us something 
about her creative identity and her willingness to operate outside the 
mainstream of what was being greenlit. If her genre mix closely mirrors the 
comparison group, it suggests her distinctiveness came from execution rather 
than subject matter. Genre composition also has financial implications,
certain genres carry higher average budgets, different audience demographics, 
and different international market performance. By using the checkboxes to 
isolate specific genres, you can ask targeted questions: how does her romantic 
comedy output compare to peers working in the same space, or how often did 
she venture into drama relative to the broader comparison set?

Column
-------------------------------------

### Genre Composition

```{r}
renderPlotly({
  p4 <- genre_tbl |>
    filter(genre %in% input$genre_filter) |>
    ggplot(aes(y = genre, fill = group,
               text = paste("Group:", group,
                            "<br>Genre:", genre))) +  # removed title — bars are aggregates
    geom_bar(position = "dodge") +
    scale_fill_manual(values = c("Meyers" = "#C9A84C", "Comparison" = "#4A2C6E")) +  # ← add this
    labs(title = "Genre Composition",
         x = "Count", y = NULL, fill = NULL) +
    theme_minimal()
  
  ggplotly(p4, tooltip = "text")
})
```

### Caption
*Compares the **genre composition** of Meyers films vs. the
comparison set. Use the checkboxes to show or hide specific genres.*



Findings
=====================================

Column 
-------------------------------------

### Our findings 

1. **ROI By Film**: Every films that Nancy Meyer was involved turned a profit. What stands out is not just how she made money but how consistent she was in doing it. Comparison films either hits extremely high return (like My Big Fat Greek Wedding at 7,274%) or land much lower.Meyers films cluster in the middle: reliable, steady, and profitable every single time. That kind of consistency is rare in Hollywood.The results also show that Meyer's films in the 1980s and 2010s are more successful to the comparison group.

2. **Budget vs. Box Office** : Almost every films sit above he break-even line, which mean nearly all of them earned more than they cost to make. Meyers films tend to have larger budgets than comparison films, yet they still land comfortably above the line. This tells us she was trusted with big money and she consistently delivered returns.

3. **Career Timeline** : We see 2 clear patterns emerged. First, once she started her career as a director in 1995 her budgets jumped significantly, studios gave her more money when she took full creative control and she always made a great return every time.Second, her IMDb ratings are surprisingly stable across 35 years, hovering between 6.5 and 7.5, which means audiences have consistently liked her work regardless of what critics said. The Intern (2015) stands out as her highest rated film on IMDb.

4. **Genre Breakdown**: Comedy shows up in almost every Meyers film ,that part is no surprise. But the more interesting finding is what comes alongside it. Meyers films are as likely to carry a Drama tag as comparison films, which suggests her movies have as much emotional weight as a typical rom-com. What id striking is that we expected Mayer to have more of the romance genre, at least as much as comparison films, yet the genre data tells a different story.This suggests that she is making something that feels romantic without fitting neatly into the Romance category in most of her films and that is what makes her stand out in her career.
