---
title: "The Meyers Effect"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: flatly
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(shiny)
library(flexdashboard)

films_tbl <- read_csv("data/films_tbl.csv", show_col_types = FALSE) |>
  mutate(decade = paste0(floor(year / 10) * 10, "s"))

genre_tbl <- read_csv("data/genre_tbl.csv", show_col_types = FALSE)
```

# About

## Row

### Background

**Who is Nancy Meyers? Why are we so invested in her?**

Nancy Meyers is one of Hollywood's most successful filmmakers — and one of its most underrated. She was born on December 8, 1949, in Philadelphia, Pennsylvania. She started her career in the 1970s writing screenplays, and went on to write, direct, and produce some of the most beloved romantic comedies of the past four decades.

Her films include **The Parent Trap** (1998), **Something's Gotta Give** (2003), **The Holiday** (2006), **It's Complicated** (2009), and **The Intern** (2015). They are funny, warm, and incredibly well-made. They also made a lot of money. Every film she has directed turned a profit,a record very few directors can claim.

### Nancy Meyers

```{r}
tags$img(src   = "nancy_mayer1.png",
         style = "width:100%; border-radius:8px;")
```

### About this project

We will looks at all 12 films Nancy Meyers was involved in between 1980 and 2015, and compares them against 14 other rom-coms from the same era. We explore four dimensions of her career

**1. Financial Performance** Do Meyers films consistently make more money relative to their budget than other rom-coms? We measure this using ROI (worldwide gross divided by production budget), to see whether her commercial success is a pattern or just luck.

**2. Budget vs. Box Office** How much did each film cost to make, and how much did it earn? We plot production budget against worldwide box office to see which films made the most money relative to what was spent. Films above the diagonal line earned more than they cost, and we want to know if Meyers films consistently land above that line compared to other rom-coms.

**3. Career Evolution** How did Nancy Meyers grow as a filmmaker since the start of her career? We trace how her budgets, box office returns, and IMDb ratings changed over time, and whether taking on the director's chair made a difference.

**4. Genre Patterns** What makes a Meyers film a Meyers film? We look at genre tags across all 26 films to see whether her work stands apart from the comparison set and whether those differences might explain why audiences see her so differently.

We chose these four dimensions because together they tell the full story of why she matters, not just as a filmmaker, but as a case study in how Hollywood measures success. Box office numbers tell us what audiences voted for with their wallets. Budgets tell us how much the studios trusted her. Ratings tell us what critics thought versus what real viewers felt. And genre patterns tell us what kind of stories she chose to tell. Taken together, they make the case that Nancy Meyers has been quietly doing something remarkable for 35 years, and that the data backs it up.

**Data sources:**

-   [Wikipedia](https://en.wikipedia.org) — financial data
-   [IMDb](https://www.imdb.com) — ratings, runtime, genres

## Row {data-height="350"}

### Her Films

```{r}
tags$img(src   = "Nancymayer.jpg",
         style = "width:100%; border-radius:8px; object-fit:cover;")
```




# ROI by Film

## Column {.sidebar}

**Filter**

```{r}
selectInput("decade_filter", "Decade",
            choices  = c("All", "1980s", "1990s", "2000s", "2010s"),
            selected = "All")

selectInput("group_filter", "Group",
            choices  = c("Both", "Meyers", "Comparison"),
            selected = "Both")
```

Films sorted by profit margin %. All films with financial data turned a profit. My Big Fat Greek Wedding (7,274% ROI) is capped at 1,400% for readability.

## Column

### Profit Margin by Film

```{r}
renderPlot({
  films_tbl |>
    filter(!is.na(profit_margin_pct)) |>
    filter(input$decade_filter == "All" | decade == input$decade_filter) |>
    filter(input$group_filter  == "Both" | group == input$group_filter) |>
    mutate(
      pct_display = pmin(profit_margin_pct, 1400),
      title       = fct_reorder(title, pct_display)
    ) |>
    ggplot(aes(x = pct_display, y = title, fill = group)) +
    geom_col() +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
    scale_fill_manual(values = c("Meyers" = "#4A2C6E", "Comparison" = "#2A7D6E")) +
    labs(x = "Profit Margin %", y = NULL, fill = NULL) +
    theme_minimal(base_size = 14)
})
```

# Budget vs Box Office

## Column {.sidebar}

**Options**

```{r}
selectInput("highlight", "Highlight a film",
            choices = c("None", sort(films_tbl$title)))
```

Bubble size = IMDb vote count. Films above the diagonal line earned more than their budget. Both axes are log-scaled.

## Column

### Budget vs. Worldwide Box Office

```{r}
renderPlot({
  films_tbl |>
    filter(!is.na(budget_usd), !is.na(box_office_usd)) |>
    mutate(label = if_else(title == input$highlight, title, NA_character_)) |>
    ggplot(aes(x = budget_usd, y = box_office_usd, size = vote_count_num, color = group)) +
    geom_abline(linetype = "dashed", color = "gray50") +
    geom_point(alpha = 0.7) +
    geom_text(aes(label = label), hjust = -0.1, size = 4, show.legend = FALSE) +
    scale_x_log10(labels = scales::label_dollar(scale = 1e-6, suffix = "M")) +
    scale_y_log10(labels = scales::label_dollar(scale = 1e-6, suffix = "M")) +
    scale_color_manual(values = c("Meyers" = "#4A2C6E", "Comparison" = "#2A7D6E")) +
    labs(x = "Production Budget", y = "Worldwide Box Office",
         color = NULL, size = "Vote count") +
    theme_minimal(base_size = 14)
})
```

# Career Timeline

## Column {.sidebar}

**Y Axis**

```{r}
radioButtons("timeline_y", NULL,
             choices  = c("Budget"      = "budget_usd",
                          "Box Office"  = "box_office_usd",
                          "IMDb Rating" = "imdb_rating"),
             selected = "budget_usd")
```

Meyers films only. Purple = directed by Meyers. Rose = writer/producer credit only.

## Column

### Nancy Meyers Career Timeline

```{r}
renderPlot({
  y_var <- input$timeline_y

  films_tbl |>
    filter(group == "Meyers", !is.na(.data[[y_var]])) |>
    mutate(role_label = if_else(meyers_directed, "Director", "Writer / Producer")) |>
    ggplot(aes(x = year, y = .data[[y_var]], color = role_label)) +
    geom_line(color = "gray80") +
    geom_point(size = 4) +
    scale_color_manual(values = c("Director" = "#4A2C6E", "Writer / Producer" = "#B5858A")) +
    scale_x_continuous(breaks = seq(1980, 2015, 5)) +
    labs(x = "Year", y = y_var, color = NULL) +
    theme_minimal(base_size = 14)
})
```

# Genre Breakdown

## Column {.sidebar}

**Filter Genres**

```{r}
checkboxGroupInput("genre_filter", NULL,
                   choices  = sort(unique(genre_tbl$genre)),
                   selected = sort(unique(genre_tbl$genre)))
```

Each film can carry multiple genre tags.

## Column

### Genre Composition: Meyers vs. Comparison

```{r}
renderPlot({
  genre_tbl |>
    filter(genre %in% input$genre_filter) |>
    ggplot(aes(y = genre, fill = group)) +
    geom_bar(position = "dodge") +
    scale_fill_manual(values = c("Meyers" = "#4A2C6E", "Comparison" = "#2A7D6E")) +
    labs(x = "Count", y = NULL, fill = NULL) +
    theme_minimal(base_size = 14)
})
```

### Conclusion

## Column {.sidebar}

1.  **ROI By Film**: Every films that Nancy Meyer was involved turned a profit. What stands out is not just how she made money but how consistent she was in doing it. Comparison films either hits extremely high return (like My Big Fat Greek Wedding at 7,274%) or land much lower.Meyers films cluster in the middle: reliable, steady, and profitable every single time. That kind of consistency is rare in Hollywood.The results also show that Meyer's films in the 1980s and 2010s are more successful to the comparison group.

2. **Budget vs. Box Office** : Almost every films sit above he break-even line, which mean nearly all of them earned more than they cost to make. Meyers films tend to have larger budgets than comparison films, yet they still land comfortably above the line. This tells us she was trusted with big money and she consistently delivered returns.

3. **Career Timeline** : We see 2 clear patterns emerged. First, once she started her career as a director in 1995 her budgets jumped significantly, studios gave her more money when she took full creative control and she always made a great return every time.Second, her IMDb ratings are surprisingly stable across 35 years, hovering between 6.5 and 7.5, which means audiences have consistently liked her work regardless of what critics said. The Intern (2015) stands out as her highest rated film on IMDb.

4.**Genre Breakdown**: Comedy shows up in almost every Meyers film ,that part is no surprise. But the more interesting finding is what comes alongside it. Meyers films are as likely to carry a Drama tag as comparison films, which suggests her movies have as much emotional weight as a typical rom-com. What id striking is that we expected Mayer to have more of the romance genre, at least as much as comparison films, yet the genre data tells a different story.This suggests that she is making something that feels romantic without fitting neatly into the Romance category in most of her films and that is what makes her stand out in her career. 